Examining background gene lists in clusterProfiler - RNA-seq example

### 2023-07-23 analysis/example_proteomics.Rmd
starting

---
title: "Examining background gene lists in clusterProfiler"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

@@

For this guide I will be using bulk RNA-seq data from a previous study, which is
deposited at NCBI GEO and SRA under accession numbers: GSE55123 and SRP038101 (Lund et al, 2014).
The experiment is designed to investigate the effect of Azacitidine treatment on AML3 cells.

The raw data have been processed by the DEE2 project, and the summary gene expression counts are
available at the dee2.io website, and programmatically with the getDEE2 bioconductor package
(Ziemann et al, 2019).

Alternatively, you could fetch data from another resource like NCBI GEO, Zenodo or from the host
storage drive.

@@

mdat <- getDEE2Metadata("hsapiens")

# get sample sheet
ss <- subset(mdat,SRP_accession=="SRP038101")

# fetch the whole set of RNA-seq data
x <- getDEE2("hsapiens",ss$SRR_accession , metadata=mdat, legacy=TRUE)
mx <- x$GeneCounts
rownames(mx) <- paste(rownames(mx),x$GeneInfo$GeneSymbol)
dim(mx)

# aza no filtering
ss$trt <- grepl("Treated",ss$Experiment_title)

ss %>%
  kbl(caption="sample sheet for Aza treatment in AML3 cells") %>%
  kable_paper("hover", full_width = F)

@@

## Data quality control

QC is important, even if you are using public transcriptome data.
For RNA-seq it is a good idea to show the number of reads for each sample.

```{r,counts1}

par(mar=c(5,7,5,1))
barplot(rev(colSums(mx)),horiz=TRUE,las=1,main="number of reads per sample in SRP038101")

Now make a MDS plot.

```{r,mds1}

mds <- cmdscale(dist(t(mx)))

# expand plot area
XMIN=min(mds[,1])*1.3
XMAX=max(mds[,1])*1.3
YMIN=min(mds[,2])*1.3
YMAX=max(mds[,2])*1.3

cols <- as.character(grepl("Treated",ss$Experiment_title))
cols <- gsub("FALSE","lightblue",cols)
cols <- gsub("TRUE","pink",cols)
plot(mds, xlab="Coordinate 1", ylab="Coordinate 2",
  xlim=c(XMIN,XMAX),ylim=c(YMIN,YMAX),
  pch=19,cex=2,col=cols, main="MDS plot")
text(cmdscale(dist(t(mx))), labels=colnames(mx) )

## Differential expression analysis

I will be using DESeq2 for DE analysis, the count matrix is prefiltered using
a detection threshold of 10 reads per sample across all samples.
All genes that meet the detection threshold will comprise the background list.
The first 6 rows of the count matrix are shows.

```{r,de1}

mxf <- mx[which(rowMeans(mx)>=10),]
dim(mxf)

head(mxf,6) %>%
  kbl(caption="Count matrix format") %>%
  kable_paper("hover", full_width = F)

dds <- DESeqDataSetFromMatrix(countData = mxf , colData = ss, design = ~ trt )
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),mxf)
def <-as.data.frame(zz[order(zz$pvalue),])

head(def,10) %>%
  kbl(caption="Top DE genes for Aza treatment") %>%
  kable_paper("hover", full_width = F)

Make a smear plot.

```{r,smearplot}

sigf <- subset(def,padj<=0.05)

DET=nrow(mxf)
NSIG=nrow(sigf)
NUP=nrow(subset(sigf,log2FoldChange>0))
NDN=nrow(subset(sigf,log2FoldChange<0))

HEADER=paste(DET,"detected genes;",NSIG,"w/FDR<0.05;",NUP,"up;",NDN,"down")

plot(log10(def$baseMean) ,def$log2FoldChange,
  cex=0.6,pch=19,col="darkgray",
  main="5-azacitidine treatment in AML3 cells",
  xlab="log10(basemean)",ylab="log2(fold change)")

points(log10(sigf$baseMean) ,sigf$log2FoldChange,
  cex=0.6,pch=19,col="red")

mtext(HEADER)

@@

@@

Get the gene sets loaded in R.

```{r,gs}
# from https://reactome.org/download/current/ReactomePathways.gmt.zip
genesets2 <- read.gmt("ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("ReactomePathways_2023-03-06.gmt")

@@

defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
defup <- unique(sapply(strsplit(defup," "),"[[",1))

defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
defdn <- unique(sapply(strsplit(defdn," "),"[[",1))

bg <- rownames(def)
bg <- unique(sapply(strsplit(bg," "),"[[",1))

# write ensembl IDs for DAVID 
writeLines(text=bg,con="ebg.txt")
writeLines(text=defup,con="eup.txt")
writeLines(text=defdn,con="edn.txt")

defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
defup <- unique(sapply(strsplit(defup," "),"[[",2))

defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
defdn <- unique(sapply(strsplit(defdn," "),"[[",2))

bg <- rownames(def)
bg <- unique(sapply(strsplit(bg," "),"[[",2))

message(paste("number of genes in background:",length(bg)))

@@

```{r,oraup}

@@

head(ora_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,oradn}

@@

head(ora_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,topbarplot2}

@@

I have also done some modification to the Reactome gene sets, to en

```{r,subsetfix}

defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
defup <- unique(sapply(strsplit(defup," "),"[[",2))

defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
defdn <- unique(sapply(strsplit(defdn," "),"[[",2))

bg <- rownames(def)
bg <- unique(sapply(strsplit(bg," "),"[[",2))
message(paste("number of genes in background:",length(bg)))

@@

```{r,gs2fix}

@@

```{r,oraupfix}

@@

head(orafix_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,oradnfix}

@@

head(orafix_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,topbarplot2fix}

@@

## Enrichment analysis with functional class scoring

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | FGSEA (check version at foot of report) |
| Statistical test used | Kolmogorov-Smirnov test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 |

```{r,fgsea}

def2 <- def
def2$genename <- sapply(strsplit(rownames(def2)," "),"[[",2)
def2 <- def2[,c("stat","genename")]
def2 <- aggregate(. ~ genename, def2, sum)
stat <- def2$stat
names(stat) <- def2$genename
fgseaRes <- fgsea(pathways=gsets, stats=stat, minSize=10, nPermSimple = 10000)
fgseaRes <- fgseaRes[order(fgseaRes$pval),]

```{r,fgsea_sig}

fgsea_up <- subset(fgseaRes,padj<0.05 & ES>0)
fgsea_ups <- fgsea_up$pathway
fgsea_dn <- subset(fgseaRes,padj<0.05 & ES<0)
fgsea_dns <- fgsea_dn$pathway

fgsea_up <-  data.frame(fgsea_up[order(-fgsea_up$ES),])

fgsea_dn <-  data.frame(fgsea_dn[order(fgsea_dn$ES),])

head(fgsea_up,n_pw) %>%
  kbl(row.names = FALSE, caption="FGSEA:Top upregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

head(fgsea_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="FGSEA:Top downregulated pathways in Aza treatment") %>%
  kable_paper("hover", full_width = F)

fgsea_up <- head(fgsea_up,n_pw)
fgsea_up_vec <- fgsea_up$ES
names(fgsea_up_vec) <- fgsea_up$pathway
fgsea_up_vec <- rev(fgsea_up_vec)

fgsea_dn <- head(fgsea_dn,n_pw)
fgsea_dn_vec <- fgsea_dn$ES * -1
names(fgsea_dn_vec) <- fgsea_dn$pathway

```{r,topbarplot3}

barplot(c(fgsea_dn_vec,fgsea_up_vec),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE Reactomes",
  xlab="ES",
  cex.main=0.9)

mtext("FCS test")

@@

v1 <- list("ORA up"=ora_ups, "ORA fix up"=orafix_ups, "FCS up"=fgsea_ups)
v2 <- list("ORA dn"=ora_dns, "ORA fix dn"=orafix_dns, "FCS dn"=fgsea_dns)
v3 <- list("ORA"=union(ora_ups,ora_dns),
  "ORA fix"=union(orafix_ups,orafix_dns),
  "FCS"=union(fgsea_ups,fgsea_dns))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated genes")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated genes")

plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated genes")

Jaccard index comparing ORA and FCS.

ora <- c(ora_ups,ora_dns)
fcs <- c(fgsea_ups,fgsea_dns)

jaccard(ora,fcs)

@@

Jaccard index comparing ORA fix and FCS.

```{r,jaccard3}

orafix <- c(orafix_ups,orafix_dns)
fcs <- c(fgsea_ups,fgsea_dns)

jaccard(orafix,fcs)

## Drill in to a specific pathway

In this case the top upregulated and downregulated pathways

```{r,eplot}

upname <- head(fgsea_up,1)$pathway
plotEnrichment(gsets[[upname]], stat, gseaParam = 1, ticksSize = 0.2)

dnname <- head(fgsea_dn,1)$pathway
plotEnrichment(gsets[[dnname]], stat, gseaParam = 1, ticksSize = 0.2)

Now a heatmap of these gene sets.

```{r,heat}

# reads per million normalisation
rpm <- apply(mxf,2, function(x) { x/sum(x) * 1000000} )
colnames(rpm) <- c("UNTR1","UNTR2","UNTR3","TRT1","TRT2","TRT3")
gnames_up <- gsets[[which(names(gsets) == upname)]]
gnames_dn <- gsets[[which(names(gsets) == dnname)]]
gene_ids <- rownames(mxf)
gene_names <- sapply(strsplit(gene_ids," "),"[[",2)
rpm_up <- rpm[which(gene_names %in% gnames_up),]
rownames(rpm_up) <- sapply(strsplit(rownames(rpm_up)," "),"[[",2)
rpm_dn <- rpm[which(gene_names %in% gnames_dn),]
rownames(rpm_dn) <- sapply(strsplit(rownames(rpm_dn)," "),"[[",2)
colsidecols <- c("blue","blue","blue","red","red","red")
heatmap.2(rpm_up,scale="row",trace="none",margins=c(10,15),main=upname,ColSideColors=colsidecols)
heatmap.2(rpm_dn,scale="row",trace="none",margins=c(10,15),main=dnname,ColSideColors=colsidecols)

>>>>

---
title: "Examining background gene lists in clusterProfiler - proteomics example"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

@@

For this report I will be using proteomics data from Mulvey et al 2021 (PMID: 34599159),
who looked at the effect of LPS on THP-1 cell lines.
Although in this paper they looked at numerous timepoints, I will be looking only at the 24hr time
point where the number of significantly differentially expressed proteins was the greatest.
The data were obtained from Supplementary Table 1 and saved to the /data folder.

@@

def <- read.table("../data/lps_24h.tsv",sep="\t",header=TRUE)

@@

# KEGG pathway analysis

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 |

@@

Get the gene sets loaded in R.
These are KEGG gene sets

```{r,gs1}
# from https://www.gsea-msigdb.org/gsea/msigdb/human/collections.jsp 16th June 2023
genesets2 <- read.gmt("../ref/c2.cp.kegg.v2023.1.Hs.symbols.gmt")
gsets <- gmtPathways("../ref/c2.cp.kegg.v2023.1.Hs.symbols.gmt")

@@

defup <- unique(subset(def,adj.P.Val<=0.05 & logFC>0)$GN)

defdn <- unique(subset(def,adj.P.Val<=0.05 & logFC<0)$GN)

bg <- unique(def$GN)

message(paste("number of proteins in background:",length(bg)))

@@

```{r,oraup1}

@@

head(ora_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,oradn1}

@@

head(ora_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,topbarplot1}

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE KEGGs",
  xlab="ES",
  cex.main=0.9)

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 + BUG FIX |

I have also done some modification to the KEGG gene sets, to en

```{r,gs2fix1}

```{r,oraupfix1}

head(orafix_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

```{r,oradnfix1}

head(orafix_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

```{r,topbarplot2fix1}

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE KEGGs",
  xlab="ES",
  cex.main=0.9)

v1 <- list("ORA up"=ora_ups, "ORA fix up"=orafix_ups)
v2 <- list("ORA dn"=ora_dns, "ORA fix dn"=orafix_dns)
v3 <- list("ORA"=union(ora_ups,ora_dns),
  "ORA fix"=union(orafix_ups,orafix_dns))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated KEGG pathways")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated KEGG pathways")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated KEGG pathways")

List gene sets identified only in the "fixed" analysis.

```{r,examine1}

ora <- union(ora_ups,ora_dns)

orafix <- union(orafix_ups,orafix_dns)

setdiff(orafix,ora)

# Reactome Gene sets

Get the gene sets loaded in R.
These are Reactome gene sets

```{r,gs2}

genesets2 <- read.gmt("../ref/ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("../ref/ReactomePathways_2023-03-06.gmt")

```{r,oraup2}

head(ora_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

```{r,oradn2}

head(ora_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

```{r,topbarplot22}

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 + BUG FIX |

@@

```{r,gs2fix2}

@@

```{r,oraupfix2}

@@

head(orafix_up) %>%
  kbl(row.names = FALSE, caption="Top upregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,oradnfix2}

@@

head(orafix_dn,n_pw) %>%
  kbl(row.names = FALSE, caption="Top downregulated pathways in LPS treatment") %>%
  kable_paper("hover", full_width = F)

@@

```{r,topbarplot2fix2}

@@

@@

```{r,venn2,fig.width=8,fig.height=5}

v1 <- list("ORA up"=ora_ups, "ORA fix up"=orafix_ups)
v2 <- list("ORA dn"=ora_dns, "ORA fix dn"=orafix_dns)
v3 <- list("ORA"=union(ora_ups,ora_dns),
  "ORA fix"=union(orafix_ups,orafix_dns))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated Reactomes")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated Reactomes")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated Reactomes")

@@

List gene sets identified only in the "fixed" analysis.

```{r,examine2}

ora <- union(ora_ups,ora_dns)

orafix <- union(orafix_ups,orafix_dns)

setdiff(orafix,ora)

### 2023-07-23 analysis/example_rnaseq.Rmd
starting

---
title: "Examining background gene lists in clusterProfiler"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

@@

@@

defup <- rownames(subset(def,padj<=0.05 & log2FoldChange>0))
defup <- unique(sapply(strsplit(defup," "),"[[",1))

defdn <- rownames(subset(def,padj<=0.05 & log2FoldChange<0))
defdn <- unique(sapply(strsplit(defdn," "),"[[",1))

bg <- rownames(def)
bg <- unique(sapply(strsplit(bg," "),"[[",1))

# write ensembl IDs for DAVID 
writeLines(text=bg,con="ebg.txt")
writeLines(text=defup,con="eup.txt")
writeLines(text=defdn,con="edn.txt")

@@

```{r,topbarplot2}

@@

@@

@@

I have also done some modification to the Reactome gene sets, to en

@@

```{r,subsetfix}

@@

```{r,gs2fix}

@@

```{r,oraupfix}

@@

```{r,oradnfix}

>>>>

---
title: "Examining background gene lists in clusterProfiler - RNA-seq example"
author: "Mark Ziemann & Anusuiya Bora"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

@@

```{r,writedata}

write.table(def,file="../data/aza_rnaseq.tsv")

# KEGG pathway analysis

@@

@@

```{r,topbarplot1}

@@

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE KEGGs",
  xlab="ES",
  cex.main=0.9)

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 + BUG FIX |

@@

I have also done some modification to the KEGG gene sets, to en

@@

```{r,subsetfix1}

@@

```{r,gs2fix1}

@@

```{r,oraupfix1}

@@

```{r,oradnfix1}

```{r,topbarplot2fix1}

barplot(c(topdn2,topup2),
  horiz=TRUE,las=1,cex.names=0.65,col=cols,
  main="top DE KEGGs",
  xlab="ES",
  cex.main=0.9)

v1 <- list("ORA up"=ora_ups, "ORA fix up"=orafix_ups)
v2 <- list("ORA dn"=ora_dns, "ORA fix dn"=orafix_dns)
v3 <- list("ORA"=union(ora_ups,ora_dns),
  "ORA fix"=union(orafix_ups,orafix_dns))

par(mar=c(10,10,10,10))
par(mfrow=c(2,1))
plot(euler(v1),quantities = list(cex = 2), labels = list(cex = 2),main="upregulated KEGG pathways")
plot(euler(v2),quantities = list(cex = 2), labels = list(cex = 2),main="downregulated KEGG pathways")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated KEGG pathways")

List gene sets identified only in the "fixed" analysis.

```{r,examine1}

ora <- union(ora_ups,ora_dns)

orafix <- union(orafix_ups,orafix_dns)

setdiff(orafix,ora)

# Reactome Gene sets

Get the gene sets loaded in R.
These are Reactome gene sets

```{r,gs2}

genesets2 <- read.gmt("../ref/ReactomePathways_2023-03-06.gmt")
gsets <- gmtPathways("../ref/ReactomePathways_2023-03-06.gmt")

```{r,oraup2}

```{r,oradn2}

```{r,topbarplot22}

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | via DEE2 at accession SRP038101 (human) |
| Other parameters | Min gene set size of 10 + BUG FIX |

```{r,gs2fix2}

```{r,oraupfix2}

```{r,oradnfix2}
